Tropical Andes Functional diversity for plants and Frugivores

### 2024-06-19 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
add new input files for all spatial grains, clean up code to reduce number of lines for 100 km fdis calculations for plants and frugivores

---
title: "Tropical Andes Functional diversity for plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(mFD)
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
library(rlang)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

@@

# Frugivores
## Create trait type table
```{r}
trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
trait_type <- c("N", "Q", "Q", "Q")
frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

## Summary of the assemblages * species dataframe
```{r}
asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final_100km)
```

## Species traits summary
```{r}
frugivore_traits_summ <- mFD::sp.tr.summary(
  tr_cat     = frug_trait_cat,   
  sp_tr      = frugivore_traits_df_final, 
  stop_if_NA = TRUE)
```

## Estimate functional trait-based distances between species
```{r}
sp_dist_frugivore <- mFD::funct.dist(
  sp_tr         = frugivore_traits_df_final,
  tr_cat        = frug_trait_cat,
  metric        = "gower",
  scale_euclid  = "scale_center",
  ordinal_var   = "classic",
  weight_type   = "equal",
  stop_if_NA    = TRUE)
```

## Generate a multidimensional space
```{r}
fspaces_quality_frugivore <- mFD::quality.fspaces(
  sp_dist             = sp_dist_frugivore,
  maxdim_pcoa         = 10,
  deviation_weighting = "absolute",
  fdist_scaling       = FALSE,
  fdendro             = "average")
```

```{r}
# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
round(fspaces_quality_frugivore$"quality_fspaces", 3)     
# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
mFD::quality.fspaces.plot(
  fspaces_quality            = fspaces_quality_frugivore,
  quality_metric             = "mad",
  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
                                 "pcoa_5d"),
  name_file                  = NULL,
  range_dist                 = NULL,
  range_dev                  = NULL,
  range_qdev                 = NULL,
  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
  gradient_deviation_quality = c(low = "yellow", high = "red"),
  x_lab                      = "Trait-based distance")
```

@@

```{r}
# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups

# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
frugivore_tr_faxes <- mFD::traits.faxes.cor(
  sp_tr          = frugivore_traits_df_final, 
  sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")], 
  plot           = TRUE)

@@

big_plot_frugivore <- mFD::funct.space.plot(
  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
  faxes           = c("PC1", "PC2", "PC3", "PC4"),
  name_file       = NULL,
  faxes_nm        = NULL,
  range_faxes     = c(NA, NA),
  color_bg        = "grey95",
  color_pool      = "darkgreen",
  fill_pool       = "white",
  shape_pool      = 21,
  size_pool       = 1,
  plot_ch         = TRUE,
  color_ch        = "black",
  fill_ch         = "white",
  alpha_ch        = 0.5,
  plot_vertices   = TRUE,
  color_vert      = "blueviolet",
  fill_vert       = "blueviolet",
  shape_vert      = 6,
  size_vert       = 1,
  plot_sp_nm      = NULL,
  nm_size         = 3,
  nm_color        = "black",
  nm_fontface     = "plain",
  check_input     = TRUE)
```

```{r}
big_plot_frugivore
```

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_frugivore <- rowSums(PAM_frugivore_site_final_100km)
subset_matrix_frugivore <- PAM_frugivore_site_final_100km[row_sums_frugivore >= 4, ]
```

@@

```{r}
#computing FD
# The number of species per assemblage has to be higher or equal to the number of traits

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
  asb_sp_w         = subset_matrix_frugivore,
  ind_vect         = "fdis",
  #did not run fdiv or fric (slows it down); c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe")
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

@@

```{r}
# Generate coordinates
subset_coords_frugivore <-site_loc_key_frugivore_100km[rowSums(PAM_frugivore_site_final_100km) >= 4, ]
subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

@@

```{r}
frugivoregridFDisTA_100km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12))
frugivoregridFDisTA_100km
```

@@

# Plants
## Create trait type table
```{r}
trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
trait_type <- c("N", "Q", "Q", "N") 
plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

## Summary of the assemblages * species dataframe
```{r}
asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
```

## Species traits summary
```{r}
plant_traits_summ <- mFD::sp.tr.summary(
  tr_cat     = plant_trait_cat,   
  sp_tr      = plant_traits_df_final, 
  stop_if_NA = TRUE)
```

```{r}
save(plant_traits_df_final, plant_trait_cat, file = "../../plant_funct.dist_files.rdata")
```

## Estimate functional trait-based distances between species
```{r}
sp_dist_plants <- mFD::funct.dist(
  sp_tr         = plant_traits_df_final,
  tr_cat        = plant_trait_cat,
  metric        = "gower",
  scale_euclid  = "scale_center",
  ordinal_var   = "classic",
  weight_type   = "equal",
  stop_if_NA    = TRUE)
```

## Generate a multidimensional space
```{r}
fspaces_quality_plants <- mFD::quality.fspaces(
  sp_dist             = sp_dist_plants,
  maxdim_pcoa         = 10,
  deviation_weighting = "absolute",
  fdist_scaling       = FALSE,
  fdendro             = "average")
```

```{r}
# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
round(fspaces_quality_plants$"quality_fspaces", 3)            
# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
mFD::quality.fspaces.plot(
  fspaces_quality            = fspaces_quality_plants,
  quality_metric             = "mad",
  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
                                 "pcoa_4d"),
  name_file                  = NULL,
  range_dist                 = NULL,
  range_dev                  = NULL,
  range_qdev                 = NULL,
  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
  gradient_deviation_quality = c(low = "yellow", high = "red"),
  x_lab                      = "Trait-based distance")
```

```{r}
#testing correlation between functional axes and traits
sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
```

```{r}
# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups

# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
plant_tr_faxes <- mFD::traits.faxes.cor(
  sp_tr          = plant_traits_df_final, 
  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
  plot           = TRUE)

@@

```{r}
#plotting functional space
sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"

big_plot_plants <- mFD::funct.space.plot(
  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
  faxes           = c("PC1", "PC2", "PC3", "PC4"),
  name_file       = NULL,
  faxes_nm        = NULL,
  range_faxes     = c(NA, NA),
  color_bg        = "grey95",
  color_pool      = "darkgreen",
  fill_pool       = "white",
  shape_pool      = 21,
  size_pool       = 1,
  plot_ch         = TRUE,
  color_ch        = "black",
  fill_ch         = "white",
  alpha_ch        = 0.5,
  plot_vertices   = TRUE,
  color_vert      = "blueviolet",
  fill_vert       = "blueviolet",
  shape_vert      = 6,
  size_vert       = 1,
  plot_sp_nm      = NULL,
  nm_size         = 3,
  nm_color        = "black",
  nm_fontface     = "plain",
  check_input     = TRUE)

#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_plant <- rowSums(PAM_plant_site_final)
subset_matrix_plant <- PAM_plant_site_final[row_sums_plant >= 4, ]
```

```{r}
big_plot_plants
```

```{r}
# getting sp_faxes_coord_plants and subset_matrix_plants names to match
sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
## check number of species names
nrow(sp_faxes_coord_plants_sub) # 54
ncol(subset_matrix_plant) #82
```

```{r}
sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)

sp_faxes_coord_plants_sub_df <- as.data.frame(sp_faxes_coord_plants_sub)
subset_matrix_plant_df <- as.data.frame(subset_matrix_plant)

subset_matrix_plant <- subset_matrix_plant_df[, which((names(subset_matrix_plant_df) %in% sp_faxes_coord_plants_sub_names)==TRUE)]

ncol(subset_matrix_plant)
subset_matrix_plant <- as.matrix(subset_matrix_plant)
```

```{r}
#computing FD
# The number of species per assemblage has to be higher or equal to the number of traits
alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_plants_sub,
  asb_sp_w         = PAM_plant_site_final,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

details_list_plant <- alpha_fd_indices_plant$"details"

#plot
plots_alpha_plant <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_plant,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

# comparing convex hulls of different hulls
plots_alpha_plant$"fric"$"patchwork"

plots_alpha_plant$"fdiv"$"patchwork"

plots_alpha_plant$"fspe"$"patchwork"

plots_alpha_plant$"fdis"$"patchwork"

plots_alpha_plant$"fori"$"patchwork"

```{r}
#Get functional dispersion
fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
```

# Extract cell values
```{r}
plantgridFDTA <- plantgridFDTA[["data"]]
frugivoregridFDTA <- frugivoregridFDTA[["data"]]
```

>>>>

---
title: "Tropical Andes Functional diversity for plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script calculates functional diversity as functional dispersion for plants and frugivores."
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(mFD)
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
library(rlang)
library(doParallel)
library(foreach)
```

# Set file paths
```{r}
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

@@

```{r}
site_loc_key_plant_75km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_75km.rds"))
site_loc_key_frugivore_75km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_75km.rds"))
PAM_plant_site_final_75km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_75km.rds"))
PAM_frugivore_site_final_75km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_75km.rds"))
```

```{r}
site_loc_key_plant_50km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_50km.rds"))
site_loc_key_frugivore_50km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_50km.rds"))
PAM_plant_site_final_50km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_50km.rds"))
PAM_frugivore_site_final_50km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_50km.rds"))
```

```{r}
site_loc_key_plant_25km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_25km.rds"))
site_loc_key_frugivore_25km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_25km.rds"))
PAM_plant_site_final_25km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_25km.rds"))
PAM_frugivore_site_final_25km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_25km.rds"))
```

```{r}
site_loc_key_plant_10km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_10km.rds"))
site_loc_key_frugivore_10km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_10km.rds"))
PAM_plant_site_final_10km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_10km.rds"))
PAM_frugivore_site_final_10km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_10km.rds"))
```

```{r}
site_loc_key_plant_5km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_5km.rds"))
site_loc_key_frugivore_5km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_5km.rds"))
PAM_plant_site_final_5km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_5km.rds"))
PAM_frugivore_site_final_5km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_5km.rds"))
```

# Create trait type tables
```{r}
trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
trait_type <- c("N", "Q", "Q", "Q")
frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

```{r}
trait_name <- c("PlantLifespan_years","GrowthForm", "SeedMass_g", "PlantHeight_m", "FruitType", "FruitLength_mm",
                "SeedWidth_mm", "FruitMass_mg", "SeedLength_mm", "DispersalSyndrome")
trait_type <- c("Q", "N", "Q", "Q", "N", "Q", "Q", "Q", "Q", "N") 
plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

```{r}
# fix Nominal traits as factor
frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)

plant_traits_df_final$GrowthForm <- as.factor(plant_traits_df_final$GrowthForm)
plant_traits_df_final$FruitType <- as.factor(plant_traits_df_final$FruitType)
plant_traits_df_final$DispersalSyndrome <- as.factor(plant_traits_df_final$DispersalSyndrome)
```

# Create function
```{r}

# Frugivores
## 100 km
```{r}
# Summary of the assemblages * species dataframe
asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final_100km)
```

```{r}
# Species traits summary
frugivore_traits_summ <- mFD::sp.tr.summary(tr_cat = frug_trait_cat, sp_tr = frugivore_traits_df_final)
```

```{r}
# Estimate functional trait-based distances between species
sp_dist_frugivore <- mFD::funct.dist( sp_tr = frugivore_traits_df_final, tr_cat = frug_trait_cat, metric = "gower",
  scale_euclid = "scale_center", ordinal_var = "classic", weight_type = "equal", stop_if_NA = TRUE)
```

```{r}
# Generate a multidimensional space
fspaces_quality_frugivore <- mFD::quality.fspaces(sp_dist = sp_dist_frugivore, maxdim_pcoa = 10, deviation_weighting = "absolute",
                                                  fdist_scaling = FALSE, fdendro = "average")
```

```{r}
# Look at the quality spaces only (MAD index looks at the mean absolute deviation from the dissimilarity matrix; want the deviation to be low meaning that the true distances have been retained in the PCA)
round(fspaces_quality_frugivore$"quality_fspaces", 3)
```

```{r}
# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 3 dimensions
mFD::quality.fspaces.plot(fspaces_quality = fspaces_quality_frugivore, quality_metric = "mad",
                          fspaces_plot = c("pcoa_3d", "pcoa_4d", "pcoa_5d"))
```

@@

```{r}
# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
frugivore_tr_faxes <- mFD::traits.faxes.cor(sp_tr = frugivore_traits_df_final, 
                                            sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
                                            plot = TRUE)

@@

big_plot_frugivore <- mFD::funct.space.plot(sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
                                            faxes = c("PC1", "PC2", "PC3", "PC4"), alpha_ch = 0.5, shape_vert = 6)
big_plot_frugivore
```

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (3)
# Calculate row sums
row_sums_frugivore <- rowSums(PAM_frugivore_site_final_100km)
subset_matrix_frugivore <- PAM_frugivore_site_final_100km[row_sums_frugivore >= 3, ]
```

@@

```{r}
# Computing FD
# The number of species per assemblage has to be higher or equal to the number of traits

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(sp_faxes_coord = sp_faxes_coord_frugivore_sub, asb_sp_w = subset_matrix_frugivore,
                                                     ind_vect = "fdis", details_returned = TRUE)

@@

### Mapping FD

```{r}
# Generate coordinates
subset_coords_frugivore <-site_loc_key_frugivore_100km[rowSums(PAM_frugivore_site_final_100km) >= 3, ]
subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

@@

```{r}
frugivoregridFDisTA_100km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    ggtitle("100 km") +
    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
frugivoregridFDisTA_100km
```

@@

# Plants

## 100 km
```{r}
# Summary of the assemblages * species dataframe
asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final_100km)
```

```{r}
# Species traits summary
plant_traits_summ <- mFD::sp.tr.summary(tr_cat = plant_trait_cat, sp_tr = plant_traits_df_final)
```

```{r}
# Estimate functional trait-based distances between species
sp_dist_plant <- mFD::funct.dist( sp_tr = plant_traits_df_final, tr_cat = plant_trait_cat, metric = "gower",
  scale_euclid = "scale_center", ordinal_var = "classic", weight_type = "equal", stop_if_NA = TRUE)
```

```{r}
# Generate a multidimensional space
fspaces_quality_plant <- mFD::quality.fspaces(sp_dist = sp_dist_plant, maxdim_pcoa = 10, deviation_weighting = "absolute",
                                                  fdist_scaling = FALSE, fdendro = "average")
```

```{r}
# Look at the quality spaces only (MAD index looks at the mean absolute deviation from the dissimilarity matrix; want the deviation to be low meaning that the true distances have been retained in the PCA)
round(fspaces_quality_plant$"quality_fspaces", 3)
```

```{r}
# Plot the quality spaces (chose to look at )
mFD::quality.fspaces.plot(fspaces_quality = fspaces_quality_plant, quality_metric = "mad",
                          fspaces_plot = c("pcoa_6d", "pcoa_7d", "pcoa_8d"))
```

```{r}
#testing correlation between functional axes and traits
sp_faxes_coord_plant <- fspaces_quality_plant$"details_fspaces"$"sp_pc_coord"
```

```{r}
# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
plant_tr_faxes <- mFD::traits.faxes.cor(sp_tr = plant_traits_df_final, 
                                            sp_faxes_coord = sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")],
                                            plot = TRUE)

@@

```{r}
#plotting functional space
sp_faxes_coord_plant <- fspaces_quality_plant$"details_fspaces"$"sp_pc_coord"

big_plot_plant <- mFD::funct.space.plot(sp_faxes_coord = sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")],
                                            faxes = c("PC1", "PC2", "PC3", "PC4"), alpha_ch = 0.5, shape_vert = 6)
big_plot_plant
```

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_plant <- rowSums(PAM_plant_site_final_100km)
subset_matrix_plant <- PAM_plant_site_final_100km[row_sums_plant >= 4, ]
```

```{r}
# match plant names
sp_faxes_coord_plant_sub <- sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")]
summary(sp_faxes_coord_plant_sub)
```

```{r}
## check number of species names
nrow(sp_faxes_coord_plant_sub)
ncol(subset_matrix_plant) 
```

```{r}
sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)
```

```{r}
sp_faxes_coord_plant_sub <- as.data.frame(sp_faxes_coord_plant_sub)
row.names(sp_faxes_coord_plant_sub) <- sp_faxes_coord_plant_sub_names

sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)

plant_names <- intersect(sp_faxes_coord_plant_sub_names, subset_matrix_plant_names)
plant_names <- na.omit(plant_names)

sp_faxes_coord_plant_sub <- sp_faxes_coord_plant_sub[ which((row.names(sp_faxes_coord_plant_sub) %in% plant_names)==TRUE), ]

subset_matrix_plant  <- as.data.frame(subset_matrix_plant)
subset_matrix_plant <- subset_matrix_plant[ ,which((colnames(subset_matrix_plant) %in% plant_names)==TRUE)]

#remove NAs
sp_faxes_coord_plant_sub <- na.omit(sp_faxes_coord_plant_sub)
subset_matrix_plant <- na.omit(subset_matrix_plant)
```

```{r}
nrow(sp_faxes_coord_plant_sub)
ncol(subset_matrix_plant)
```

```{r}
sp_faxes_coord_plant_sub <- as.matrix(sp_faxes_coord_plant_sub)
subset_matrix_plant <- as.matrix(subset_matrix_plant)
```

```{r}
# Computing FD
# The number of species per assemblage has to be higher or equal to the number of traits

alpha_fd_indices_plant <- mFD::alpha.fd.multidim(sp_faxes_coord = sp_faxes_coord_plant_sub, asb_sp_w = subset_matrix_plant,
                                                     ind_vect = "fdis", details_returned = TRUE)

details_list_plant <- alpha_fd_indices_plant$"details"
```

```{r}
# Get functional dispersion
fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
```

```{r}
# Generate coordinates
subset_coords_plant <-site_loc_key_plant_100km[rowSums(PAM_plant_site_final_100km) >= 4, ]
subset_coords_plant_sp <-subset_coords_plant[,1:2 ]

plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
```

```{r}
# Convert the dataframe to sf format
spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude", "Latitude"))
```

```{r}
# set crs of sf objects
spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(5389)
```

```{r}
# Extract coordinates 
spatial_fdis_plant_coords <- st_coordinates(spatial_fdis_plant)

# Add coordinates as separate columns
spatial_fdis_plant <- spatial_fdis_plant %>%
  mutate(x = spatial_fdis_plant_coords[, 1], y = spatial_fdis_plant_coords[, 2])
```

```{r}
summary(spatial_fdis_plant)
```

spatial_fdis_plant$cellid <- st_intersects(spatial_fdis_plant, TAGrid_100km)$nn

spatial_fdis_plant_grid <- TAGrid_100km %>%
   st_join(spatial_fdis_plant) %>%
   group_by(cellid) %>%
   summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
   ungroup()
```

```{r}
summary(spatial_fdis_plant_grid)
```

```{r}
plantgridFDisTA_100km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    ggtitle("100 km") +
    scale_fill_distiller(palette = "Greens", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
plantgridFDisTA_100km
```

```{r}
# Extract cell values
plant_cellFDis_100km <- spatial_fdis_plant_grid
hist(plant_cellFDis_100km$fdis_value)
```

## 75 km

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_plant <- rowSums(PAM_plant_site_final_75km)
subset_matrix_plant <- PAM_plant_site_final_75km[row_sums_plant >= 4, ]
```

```{r}
# match plant names
sp_faxes_coord_plant_sub <- sp_faxes_coord_plant[ , c("PC1", "PC2", "PC3", "PC4")]
summary(sp_faxes_coord_plant_sub)
```

```{r}
## check number of species names
nrow(sp_faxes_coord_plant_sub)
ncol(subset_matrix_plant) 
```

```{r}
sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)
```

```{r}
sp_faxes_coord_plant_sub <- as.data.frame(sp_faxes_coord_plant_sub)
row.names(sp_faxes_coord_plant_sub) <- sp_faxes_coord_plant_sub_names

sp_faxes_coord_plant_sub_names <- row.names(sp_faxes_coord_plant_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)

plant_names <- intersect(sp_faxes_coord_plant_sub_names, subset_matrix_plant_names)
plant_names <- na.omit(plant_names)

sp_faxes_coord_plant_sub <- sp_faxes_coord_plant_sub[ which((row.names(sp_faxes_coord_plant_sub) %in% plant_names)==TRUE), ]

subset_matrix_plant  <- as.data.frame(subset_matrix_plant)
subset_matrix_plant <- subset_matrix_plant[ ,which((colnames(subset_matrix_plant) %in% plant_names)==TRUE)]

#remove NAs
sp_faxes_coord_plant_sub <- na.omit(sp_faxes_coord_plant_sub)
subset_matrix_plant <- na.omit(subset_matrix_plant)
```

```{r}
nrow(sp_faxes_coord_plant_sub)
ncol(subset_matrix_plant)
```

```{r}
sp_faxes_coord_plant_sub <- as.matrix(sp_faxes_coord_plant_sub)
subset_matrix_plant <- as.matrix(subset_matrix_plant)
```

```{r}
# Computing FD
# The number of species per assemblage has to be higher or equal to the number of traits

alpha_fd_indices_plant <- mFD::alpha.fd.multidim(sp_faxes_coord = sp_faxes_coord_plant_sub, asb_sp_w = subset_matrix_plant,
                                                     ind_vect = "fdis", details_returned = TRUE)

details_list_plant <- alpha_fd_indices_plant$"details"
```

```{r}
# Get functional dispersion
fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
```

```{r}
# Generate coordinates
subset_coords_plant <-site_loc_key_plant_75km[rowSums(PAM_plant_site_final_75km) >= 4, ]
subset_coords_plant_sp <-subset_coords_plant[,1:2 ]

plant_fd_sp_75km <- data.frame(subset_coords_plant_sp, fdis_plant)
```

```{r}
# Convert the dataframe to sf format
spatial_fdis_plant_75km <- st_as_sf(plant_fd_sp_75km, coords = c("Longitude", "Latitude"))
```

```{r}
# set crs of sf objects
spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>% st_set_crs(5389)
```

```{r}
# Extract coordinates 
spatial_fdis_plant_coords_75km <- st_coordinates(spatial_fdis_plant_75km)

# Add coordinates as separate columns
spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>%
  mutate(x = spatial_fdis_plant_coords_75km[, 1], y = spatial_fdis_plant_coords_75km[, 2])
```

```{r}
summary(spatial_fdis_plant_75km)
```

```{r}
TAGrid__75km <- TApoly %>%
  st_make_grid(cellsize = c(75000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())

spatial_fdis_plant_75km$cellid <- st_intersects(spatial_fdis_plant_75km, TAGrid_75km)$nn

spatial_fdis_plant_grid_75km <- TAGrid_75km %>%
   st_join(spatial_fdis_plant_75km) %>%
   group_by(cellid) %>%
   summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
   ungroup()
```

```{r}
summary(spatial_fdis_plant_grid_75km)
```
